Notes on Validated Model Counting \\ Version of \today

### 2022-06-13 doc/mc-notes.tex
Expand example cost functions.  Add operator deletion

{\bf Example 1}: Let $\ring$ be the set of rational numbers of the form $a\cdot 2^b$ where $a$ and $b$ are integers.
Let $\cost(x_i) = 1/2$ for all variables $x_i$.  The cost of every
assignment is then $1/2^{n}$, and the cost of a function is its
{\em density}, denoted $\density(f)$.  That is, the density of $f$, for which
$0 \leq \density(f) \leq 1$, 
is the fraction of assignments for which $f$
evaluates to 1, with $\density(\zero) = 0$ and $\density(\one) = 1$.  The density of a function
$f$ can be scaled by $2^n$ to compute the total number of models
$|\interpset{f}|$.  This is the core task of model counting.  Using
density as the metric, rather than the number of models, has the advantage that it does not vary when the
function is embedded in a larger domain $X' \supseteq X$.

{\bf Example 2}: Let $\ring$ be the set of rational numbers.  Assign a
{\em weight} $w_i$ to each variable $x_i$ such that $0 \leq w_i \leq
1$ and let $\cost(x_i) = w_i$.  This implements {\em weighted model counting}
under the restrictions that: 
1) the
weight of an assignment equals the product of the weights of its
literals, and 2) the weight of a variable $x_i$ and its negation
$\obar{x}_i$ sum to 1.

{\bf Example 3}: Let $\ring$ be a finite field with $z = |\ring| \geq 2n$,
and let $\hashset$ be the set of functions
mapping elements of $X$ to elements of $\ring$.  For
two distinct functions $f_1$ and $f_2$ and a randomly chosen $\fhash
\in \hashset$, the probability that $h(f_1) \not = h(f_2)$ will be at least
$\left(1-\frac{1}{z}\right)^n \geq \left(1-\frac{1}{2n}\right)^n >  1/2$.  Therefore, these functions can be used as part of a
randomized algorithm for equivalence testing~\cite{blum:ipl:1980}.

@@

\begin{table}
  \caption{Recursive Definition of Separable Schemas}
  \label{tab:schema}
%%A) Basic Operations\\
\begin{center}
\begin{tabular}{cccc}
\toprule
  \makebox[15mm]{$S$}           & \makebox[30mm]{Restrictions} & \makebox[30mm]{$D(S)$} & \makebox[30mm]{$\interpset{S}$}\\
\midrule
  $\nil$  & None  & $\emptyset$    & $\emptyset$ \\
  $\tautology$        & None  & $\emptyset$    & $\uinterp$ \\
$x_i$           & None  & $\{ x_i \}$    & $\{ \interp | \interp(x_i) = \tautology \}$ \\
$\neg S_1$      & None  & $\depend(S_1)$ & $\uinterp - \interpset{S_1}$ \\
$S_1 \pand S_2$ & $\depend(S_1) \cap \depend(S_2) = \emptyset$ & $\depend(S_1) \cup \depend(S_2)$ & $\interpset{S_1} \cap \interpset{S_2}$ \\
$S_1 \por S_2$  & $\interpset{S_1} \cap \interpset{S_2} = \emptyset$ &  $\depend(S_1) \cup \depend(S_2)$ &  $\interpset{S_1} \cup \interpset{S_2}$ \\
\bottomrule
\end{tabular}
\end{center}
%% B) Composite Operations\\
%% \begin{center}
%% \begin{tabular}{cc}
%% \toprule
%%   $S$           & Definition \\
%% \midrule
%% \makebox[30mm]{$\pite(S_1, S_2, S_3)$} & \makebox[40mm]{$[S_1 \pand S_2] \por [\neg S_1 \pand S_2]$}\\
%% \bottomrule
%% \end{tabular}
%% \end{center}
\end{table}

The set of schemas over a set of variables
$\{x_1, x_2, \ldots, x_n\}$
can be defined recursively, as is shown in Table
\ref{tab:schema}.  Each schema $S$ has an associated dependency set
$\depend(S)$ and an associated set of models $\interpset{S}$.
%% We
%% divide the operations into {\em basic} ones that are fundamental to
%% the representation, and {\em composite} ones that can be constructed
%% from basic operations.  The latter category consists of just the $\pite$ operation.

A key property of a Boolean formula represented by separable schema
$S$ is that, for any separable cost function $\cost$, the cost of the
formula $\cost(S)$ can be computed with a linear number of ring
operations.

@@

\begin{table}
  \caption{CRAT Step Types.  $C$: clause identifier, $L$: literal, $V$: variable}
  \label{tab:crat:syntax}
\centering{
  \begin{tabular}{lllll}
    \multicolumn{4}{c}{Rule} & \multicolumn{1}{c}{Description} \\
    \midrule
    \makebox[5mm][l]{$C$} & \makebox[10mm][l]{\tt i}  & \makebox[15mm][l]{$L^{*}$ {\tt 0}} & \makebox[15mm][l]{}  & \makebox[20mm][l]{Input clause} \\
    $C$ & {\tt a} & $L^{*}$ {\tt 0} & $C^{+}$  {\tt 0} & Add RUP clause.  Explicit hint \\
    $C$ & {\tt a} & $L^{*}$ {\tt 0} &  {\tt * 0} & Add RUP clause.  Inferred hint \\
     & {\tt d} & $C$             & $C^{+}$  {\tt 0} & Delete RUP clause.  Explicit hint \\
     & {\tt d} & $C$             & {\tt * 0} & Delete RUP clause.  Inferred hint \\
    \midrule
    $C$    & {\tt p} & $V \; L \; L$    &                  & Declare $\pand$ operation \\
    $C$    & {\tt s} & $V \; L \; L$    & $C^{+}$ {\tt 0}  & Declare $\por$ operation.  Explicit hint \\ 
    $C$    & {\tt s} & $V \; L \; L$    & {\tt * 0}  & Declare $\por$ operation.  Inferred hint \\ 
%%        & {\tt ite} & $V \; L \; L\; L$    &                  & Declare $\pite$ operation \\
  \end{tabular}
  }
\end{table}

@@

The first set of proof rules are similar to those in other clausal
proofs.  Our syntax optionally allows input clauses to be listed with
a rule of type {\tt i}.  Clauses can be added via RUP addition
(command {\tt a}), with the sequence of antecedent clauses (the
``hint'') either provided explicitly or to be inferred (indicated by the character `{\tt *}') by the proof
checker.  Similarly for clause deletion (command {\tt d}).

@@

The declaration of a $\por$ operation has the form:
\begin{center}
\begin{tabular}{ccccccc}
  \makebox[5mm]{$i$} & \makebox[5mm]{{\tt s}} & \makebox[5mm]{$v$} & \makebox[5mm]{$l_1$} & \makebox[5mm]{$l_2$} 
\makebox[5mm]{$H$} & \makebox[5mm]{$\texttt{0}$} \\
\end{tabular}
\end{center}
where $i$ is a new clause ID, $v$ is a positive integer that does
not correspond to any previous variable, and $l_1$ and $l_2$ are
signed integers representing literals of existing variables.  Hint $H$
can consist either of the single character \texttt{*}, or it can be a
sequence of clause IDs.

This declaration implicitly causes three clauses to be added to the proof:
\begin{center}
\begin{tabular}{cccc}
\makebox[10mm]{ID} & \multicolumn{3}{c}{Clause} \\
  $i$ & $-v$ & $l_1$ & $l_2$ \\
  $i\!+\!1$ & $v$ & $-l_1$ \\
  $i\!+\!2$ & $v$ & $-l_2$ \\
\end{tabular}
\end{center}
The explicit or implied hints must provide a RUP proof of the clause $-l_1 \lor -l_2$.

@@

\begin{figure}
  A).  CRAT File Contents
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2}        & {\tt p}   & {\tt 4 -2 3}  &            & & Declare $p_4 = \obar{x}_2 \pand x_3$ \\
    {\tt 5}        & {\tt s}   & {\tt 5 2 4}   & {\tt 3 0}  & & Declare $s_5 = x_2 \por p_4$ \\
    {\tt 8}        & {\tt p}   & {\tt 6 -1 5}  &            & & Declare $p_6 = \obar{x}_1 \pand s_5$ \\
    {\tt 11}         & {\tt s}   & {\tt 7 1 6}   & {\tt 9 0}  & & Declare $s_7 = x_1 \por p_6$ \\
    {\tt 14} & {\tt a}  & {\tt 7 0} & {\tt 12 13 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt d}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
B). Proof Clauses
    \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2} &     & {\tt 4  2 -3 0}  &     & & Defining clauses for $p_4$ \\ 
    {\tt 3} &     & {\tt -4 -2 0}    & & & \\ %%Defining clause for $p_4$ \\ 
    {\tt 4} &     & {\tt -4 3 0}    &  & & \\ %%Defining clause for $p_4$ \\ 
            &     & {\tt -2 -4 0}   & {\tt 3 0}  & &  Mutual exclusion proof for $s_5$ \\
    {\tt 5} &     & {\tt -5 2 4 0}  &     & & Defining clauses for $s_5$ \\ 
    {\tt 6} &     & {\tt  5 -2 0}    &  &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 7} &     & {\tt  5 -4 0}    &  & &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 8} &     & {\tt 6  1 -5 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 9} &     & {\tt -6 -1 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
    {\tt 10} &     & {\tt -6 5 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
            &     & {\tt -1 -6 0}   & {\tt 9 0}  & &  Mutual exclusion proof for $s_7$ \\
    {\tt 11} &     & {\tt -7 1 6 0}  &  & & Defining clauses for $s_7$ \\ 
    {\tt 12} &     & {\tt  7 -1 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 13} &     & {\tt  7 -6 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 14} &     & {\tt 7 0} & {\tt 12 13 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt d}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
  \caption{CRAT file \#1 for formula $x_1 \lor x_2 \lor x_3$, and the resulting set of proof clauses}
  \label{fig:c3:crat}
\end{figure}

@@

    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} & {\tt i}   & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} & {\tt p}   & {\tt 4 -2 3}  &            & & Declare $p_4 = \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 6} & {\tt p}   & {\tt 5 -1 4}  &             & & Declare $p_5 = \obar{x}_1 \pand p_4 = \obar{x}_1 \pand \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 9} & {\tt a}  & {\tt -5 0} & {\tt 7 8 4 5 1 0} & & Assert $t_1 = \obar{p}_5 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt d}  & {\tt 1 } & {\tt 3 6 7 0} & & Delete clause $C_1$\\
    {\tt 10} & {\tt p}   & {\tt 6 1 -2} &              & & Declare $p_6 = x_1 \pand \obar{x}_2$ \\    
    {\tt 13} & {\tt a} & {\tt -6 0} & {\tt 11 12 2 0}  & & Assert $t_2 = \obar{p}_6 = \obar{x}_1 \lor x_2$ \\
            & {\tt d}  & {\tt 2 } & {\tt 10 13 0} & & Delete clause $C_2$\\
    {\tt 14} & {\tt s}   & {\tt 7 5 6}   & {\tt 7 11 0}  & & Declare $s_7 = \obar{t}_1 \por \obar{t}_2$ \\
    {\tt 17} & {\tt a}  & {\tt -7 0}       & {\tt 9 13 0} & & Assert $t_3 = \obar{s}_7 = t_1 \land t_2$ \\
             & {\tt d}  & {\tt 9}          & {\tt 15 17 0} & & Delete $t_1$\\
             & {\tt d}  & {\tt 13}         & {\tt 16 17 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
B). Proof Clauses
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} &    & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} &   & {\tt 4 2 3 0} &     & & Defining clauses for $p_4$ \\
    {\tt 4} &   & {\tt -4 -2 0} &     & & \\
    {\tt 5} &   & {\tt -4 -3 0} &     & & \\

@@

    {\tt 9} &   & {\tt -5 0} & {\tt 7 8 4 5 1 0} & & Assert $t_1 = \obar{p}_5 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt d}  & {\tt 1 } & {\tt 3 6 7 0} & & Delete clause $C_1$\\

    {\tt 10} &   & {\tt 6 -1 2 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 11} &   & {\tt -6 1 0}    &  & & \\
    {\tt 12} &   & {\tt -6 -2 0}    &  & & \\ 
    {\tt 13} &  & {\tt -6 0} & {\tt 11 12 2 0}  & & Assert $t_2 = \obar{p}_6 = \obar{x}_1 \lor x_2$ \\
            & {\tt d}  & {\tt 2 } & {\tt 10 13 0} & & Delete clause $C_2$\\
             &          & {\tt -5 -6 0} & {\tt 7 11 0}  & & Mutual exclusion proof for $s_7$ \\
    {\tt 14} &   & {\tt -7 5 6 0}  &     & & Defining clauses for $s_7$ \\ 
    {\tt 15} &   & {\tt  7 -5 0}    &  &  \\  
    {\tt 16} &   & {\tt  7 -6 0}    &  & &  \\
    {\tt 17} &   & {\tt -7 0}       & {\tt 9 13 0} & & Assert $t_3 = \obar{s}_7 = t_1 \land t_2$ \\
             & {\tt d}  & {\tt 9}          & {\tt 15 17 0} & & Delete $t_1$\\
             & {\tt d}  & {\tt 13}         & {\tt 16 17 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
  \caption{CRAT file \#2A for formula $\phi_2 = (x_1 \lor x_2 \lor x_3) \land (\obar{x}_1 \lor x_2)$, and the resulting set of proof clauses}
  \label{fig:p2-bdd:crat}
\end{figure}

>>>>

{\bf Example 1a}: Let $\ring$ be the set of rational numbers of the form $a\cdot 2^b$ where $a$ and $b$ are integers.
Let $\cost(x_i) = 1/2$ for all variables $x_i$.  The cost of every
assignment is then $1/2^{n}$, and the cost of a function is its
{\em density}, denoted $\density(f)$.  That is, the density of $f$, for which
$0 \leq \density(f) \leq 1$, 
is the fraction of assignments for which $f$
evaluates to 1, with $\density(\zero) = 0$ and $\density(\one) = 1$.  The density of a function
$f$ can be scaled by $2^n$ to compute the total number of models
$|\interpset{f}|$.  This is the core task of model counting.  Using
density as the metric, rather than the number of models, has the advantage that it does not vary when the
function is embedded in a larger domain $X' \supseteq X$.

{\bf Example 1b}: Let $\ring$ be the set of rational numbers.  Assign a
{\em weight} $w(x_i)$ to each variable $x_i$ such that
$0 \leq w(x_i) \leq 1$
and let $\cost(x_i) = w(x_i)$.  This implements {\em weighted model counting}
under the restrictions that: 
1) the
weight of an assignment equals the product of the weights of its
literals, and 2) the weight of a variable $x_i$ and its negation
$\obar{x}_i$ sum to 1.

{\bf Example 1c}: Let $\ring$ be the set of rational numbers.  Assign
{\em weights} to each literal.  In particular, for each variable
$x_i$, define weights $w(x_i)$ and $w(\obar{x}_i)$, such that
$\beta(x_i) \doteq w(x_i) + w(\obar{x}_i) \not = 0$.  We can convert
this to the formulation of Example 1b by using weight $w'(x_i) =
w(x_i)/\beta(x_i)$ and then scaling the final result by $\prod_{1 \leq
  i \leq n} \beta(x_i)$.  Scaling eliminates the need to have the weight
of each variable and its negation sum to 1.  This form is sufficiently general
to handle the benchmarks used in weighted model counting track of the Model Counting Competition.
Using $w(x_i) = w(\obar{x}_i) = 1$ for each variable $x_i$,
gives the standard model counting of Example 1a.

{\bf Example 2}: Let $\ring$ be a finite field with $z = |\ring| \geq 2n$,
and let $\hashset$ be the set of functions
mapping elements of $X$ to elements of $\ring$.  For
two distinct functions $f_1$ and $f_2$ and a randomly chosen $\fhash
\in \hashset$, the probability that $h(f_1) \not = h(f_2)$ will be at least
$\left(1-\frac{1}{z}\right)^n \geq \left(1-\frac{1}{2n}\right)^n >  1/2$.  Therefore, these functions can be used as part of a
randomized algorithm for equivalence testing~\cite{blum:ipl:1980}.

@@

\begin{table}
  \caption{Recursive Definition of Separable Schemas and their Costs}
  \label{tab:schema}
%%A) Basic Operations\\
\begin{center}
\begin{tabular}{ccccc}
\toprule
  \makebox[15mm]{$S$}  & \makebox[30mm]{Restrictions} & \makebox[30mm]{$D(S)$} & \makebox[30mm]{$\interpset{S}$} & \makebox[25mm]{$\cost(S)$}\\
\midrule
  $\nil$  & None  & $\emptyset$    & $\emptyset$  & $0$\\
  $\tautology$        & None  & $\emptyset$    & $\uinterp$ & $1$ \\
$x_i$           & None  & $\{ x_i \}$    & $\{ \interp | \interp(x_i) = \tautology \}$ & $\cost(x_i)$ \\
$\neg S_1$      & None  & $\depend(S_1)$ & $\uinterp - \interpset{S_1}$ & $\oneminus\cost(S)$ \\
$S_1 \pand S_2$ & $\depend(S_1) \cap \depend(S_2) = \emptyset$ & $\depend(S_1) \cup \depend(S_2)$ & $\interpset{S_1} \cap \interpset{S_2}$
& $\cost(S_1) \cdot \cost(S_2)$\\
$S_1 \por S_2$  & $\interpset{S_1} \cap \interpset{S_2} = \emptyset$ &  $\depend(S_1) \cup \depend(S_2)$ &  $\interpset{S_1} \cup \interpset{S_2}$
& $\cost(S_1) + \cost(S_2)$\\
\bottomrule
\end{tabular}
\end{center}
%% B) Composite Operations\\
%% \begin{center}
%% \begin{tabular}{cc}
%% \toprule
%%   $S$           & Definition \\
%% \midrule
%% \makebox[30mm]{$\pite(S_1, S_2, S_3)$} & \makebox[40mm]{$[S_1 \pand S_2] \por [\neg S_1 \pand S_2]$}\\
%% \bottomrule
%% \end{tabular}
%% \end{center}
\end{table}

The set of schemas over a set of variables
$\{x_1, x_2, \ldots, x_n\}$
can be defined recursively, as is shown in Table
\ref{tab:schema}.  Each schema $S$ has an associated dependency set
$\depend(S)$ and an associated set of models $\interpset{S}$.
%% We
%% divide the operations into {\em basic} ones that are fundamental to
%% the representation, and {\em composite} ones that can be constructed
%% from basic operations.  The latter category consists of just the $\pite$ operation.
A key property of a Boolean formula represented by separable schema
$S$ is that, for any separable cost function $\cost$, the cost of the
formula $\cost(S)$ can be computed with a linear number of ring
operations, as is shown in the final column of Table~\ref{tab:schema}.

@@

\begin{table}
  \caption{CRAT Step Types.  $C$: clause identifier, $L$: literal, $V$: variable}
  \label{tab:crat:syntax}
\centering{
  \begin{tabular}{lllll}
    \multicolumn{4}{c}{Rule} & \multicolumn{1}{c}{Description} \\
    \midrule
    \makebox[5mm][l]{$C$} & \makebox[10mm][l]{\tt i}  & \makebox[15mm][l]{$L^{*}$ {\tt 0}} & \makebox[15mm][l]{}  & \makebox[20mm][l]{Input clause} \\
    $C$ & {\tt a} & $L^{*}$ {\tt 0} & $C^{+}$  {\tt 0} & Add RUP clause.  Explicit hint \\
    $C$ & {\tt a} & $L^{*}$ {\tt 0} &  {\tt * 0} & Add RUP clause.  Inferred hint \\
     & {\tt dc} & $C$             & $C^{+}$  {\tt 0} & Delete RUP clause.  Explicit hint \\
     & {\tt dc} & $C$             & {\tt * 0} & Delete RUP clause.  Inferred hint \\
    \midrule
    $C$    & {\tt p} & $V \; L \; L$    &                  & Declare $\pand$ operation \\
    $C$    & {\tt s} & $V \; L \; L$    & $C^{+}$ {\tt 0}  & Declare $\por$ operation.  Explicit hint \\ 
    $C$    & {\tt s} & $V \; L \; L$    & {\tt * 0}  & Declare $\por$ operation.  Inferred hint \\ 
%%        & {\tt ite} & $V \; L \; L\; L$    &                  & Declare $\pite$ operation \\
     & {\tt do} & $V$             &  & Delete operation. \\

  \end{tabular}
  }
\end{table}

@@

The first set of proof rules are similar to those in other clausal
proofs.  Our syntax optionally allows input clauses to be listed with
a rule of type {\tt i}.  Clauses can be added via RUP addition
(command {\tt a}), with the sequence of antecedent clauses (the
``hint'') either provided explicitly or to be inferred (indicated by the character `{\tt *}') by the proof
checker.  Similarly for clause deletion (command {\tt dc}).

@@

The declaration of a $\por$ operation has the form:
\begin{center}
\begin{tabular}{ccccccc}
  \makebox[5mm]{$i$} & \makebox[5mm]{{\tt s}} & \makebox[5mm]{$v$} & \makebox[5mm]{$l_1$} & \makebox[5mm]{$l_2$} 
\makebox[5mm]{$H$} & \makebox[5mm]{$\texttt{0}$} \\
\end{tabular}
\end{center}
where $i$ is a new clause ID, $v$ is a positive integer that does
not correspond to any previous variable, and $l_1$ and $l_2$ are
signed integers representing literals of existing variables.  Hint $H$
can consist either of the single character \texttt{*}, or it can be a
sequence of clause IDs.
This declaration implicitly causes three clauses to be added to the proof:
\begin{center}
\begin{tabular}{cccc}
\makebox[10mm]{ID} & \multicolumn{3}{c}{Clause} \\
  $i$ & $-v$ & $l_1$ & $l_2$ \\
  $i\!+\!1$ & $v$ & $-l_1$ \\
  $i\!+\!2$ & $v$ & $-l_2$ \\
\end{tabular}
\end{center}
The explicit or implied hints must provide a RUP proof of the clause $-l_1 \lor -l_2$.

Finally, operator deletion ({\tt do}) specifies which operator to
delete via the extension variable associated with the operator.  Doing
so for operator $v$ requires 1) there is no (undeleted) operator $u$
having a literal of $v$ as one of its inputs, and 2) there is no
(undeleted) RUP clause containing a literal of $v$.

@@

\begin{figure}
  A).  CRAT File Contents
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2}        & {\tt p}   & {\tt 4 -2 3}  &            & & Declare $p_4 = \obar{x}_2 \pand x_3$ \\
    {\tt 5}        & {\tt s}   & {\tt 5 2 4}   & {\tt 3 0}  & & Declare $s_5 = x_2 \por p_4$ \\
    {\tt 8}        & {\tt p}   & {\tt 6 -1 5}  &            & & Declare $p_6 = \obar{x}_1 \pand s_5$ \\
    {\tt 11}         & {\tt s}   & {\tt 7 1 6}   & {\tt 9 0}  & & Declare $s_7 = x_1 \por p_6$ \\
    {\tt 14} & {\tt a}  & {\tt 7 0} & {\tt 12 13 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
B). Proof Clauses
    \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2} &     & {\tt 4  2 -3 0}  &     & & Defining clauses for $p_4$ \\ 
    {\tt 3} &     & {\tt -4 -2 0}    & & & \\ %%Defining clause for $p_4$ \\ 
    {\tt 4} &     & {\tt -4 3 0}    &  & & \\ %%Defining clause for $p_4$ \\ 
            &     & {\tt -2 -4 0}   & {\tt 3 0}  & &  Mutual exclusion proof for $s_5$ \\
    {\tt 5} &     & {\tt -5 2 4 0}  &     & & Defining clauses for $s_5$ \\ 
    {\tt 6} &     & {\tt  5 -2 0}    &  &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 7} &     & {\tt  5 -4 0}    &  & &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 8} &     & {\tt 6  1 -5 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 9} &     & {\tt -6 -1 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
    {\tt 10} &     & {\tt -6 5 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
            &     & {\tt -1 -6 0}   & {\tt 9 0}  & &  Mutual exclusion proof for $s_7$ \\
    {\tt 11} &     & {\tt -7 1 6 0}  &  & & Defining clauses for $s_7$ \\ 
    {\tt 12} &     & {\tt  7 -1 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 13} &     & {\tt  7 -6 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 14} &     & {\tt 7 0} & {\tt 12 13 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
  \caption{CRAT file \#1 for formula $x_1 \lor x_2 \lor x_3$, and the resulting set of proof clauses}
  \label{fig:c3:crat}
\end{figure}

@@

    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} & {\tt i}   & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} & {\tt p}   & {\tt 4 -2 3}  &            & & Declare $p_4 = \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 6} & {\tt p}   & {\tt 5 -1 4}  &             & & Declare $p_5 = \obar{x}_1 \pand p_4 = \obar{x}_1 \pand \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 9} & {\tt a}  & {\tt -5 0} & {\tt 7 8 4 5 1 0} & & Assert $t_1 = \obar{p}_5 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 6 7 0} & & Delete clause $C_1$\\
    {\tt 10} & {\tt p}   & {\tt 6 1 -2} &              & & Declare $p_6 = x_1 \pand \obar{x}_2$ \\    
    {\tt 13} & {\tt a} & {\tt -6 0} & {\tt 11 12 2 0}  & & Assert $t_2 = \obar{p}_6 = \obar{x}_1 \lor x_2$ \\
            & {\tt dc}  & {\tt 2 } & {\tt 10 13 0} & & Delete clause $C_2$\\
    {\tt 14} & {\tt s}   & {\tt 7 5 6}   & {\tt 7 11 0}  & & Declare $s_7 = \obar{t}_1 \por \obar{t}_2$ \\
    {\tt 17} & {\tt a}  & {\tt -7 0}       & {\tt 9 13 0} & & Assert $t_3 = \obar{s}_7 = t_1 \land t_2$ \\
             & {\tt dc}  & {\tt 9}          & {\tt 15 17 0} & & Delete $t_1$\\
             & {\tt dc}  & {\tt 13}         & {\tt 16 17 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
B). Proof Clauses
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} &    & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} &   & {\tt 4 2 3 0} &     & & Defining clauses for $p_4$ \\
    {\tt 4} &   & {\tt -4 -2 0} &     & & \\
    {\tt 5} &   & {\tt -4 -3 0} &     & & \\

@@

    {\tt 9} &   & {\tt -5 0} & {\tt 7 8 4 5 1 0} & & Assert $t_1 = \obar{p}_5 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 6 7 0} & & Delete clause $C_1$\\

    {\tt 10} &   & {\tt 6 -1 2 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 11} &   & {\tt -6 1 0}    &  & & \\
    {\tt 12} &   & {\tt -6 -2 0}    &  & & \\ 
    {\tt 13} &  & {\tt -6 0} & {\tt 11 12 2 0}  & & Assert $t_2 = \obar{p}_6 = \obar{x}_1 \lor x_2$ \\
            & {\tt dc}  & {\tt 2 } & {\tt 10 13 0} & & Delete clause $C_2$\\
             &          & {\tt -5 -6 0} & {\tt 7 11 0}  & & Mutual exclusion proof for $s_7$ \\
    {\tt 14} &   & {\tt -7 5 6 0}  &     & & Defining clauses for $s_7$ \\ 
    {\tt 15} &   & {\tt  7 -5 0}    &  &  \\  
    {\tt 16} &   & {\tt  7 -6 0}    &  & &  \\
    {\tt 17} &   & {\tt -7 0}       & {\tt 9 13 0} & & Assert $t_3 = \obar{s}_7 = t_1 \land t_2$ \\
             & {\tt dc}  & {\tt 9}          & {\tt 15 17 0} & & Delete $t_1$\\
             & {\tt dc}  & {\tt 13}         & {\tt 16 17 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
  \caption{CRAT file \#2A for formula $\phi_2 = (x_1 \lor x_2 \lor x_3) \land (\obar{x}_1 \lor x_2)$, and the resulting set of proof clauses}
  \label{fig:p2-bdd:crat}
\end{figure}
